Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-04-08 paper.tex
Clarified blank/bye, removed "he", deleted "... omit minor details ...", creared up "balance"

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e.\ \emph{skippers}, compete against each other in a series of matches taking
place in rounds, known as \emph{flights}.  A match is composed of two skippers, with each skipper in
a boat on their own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the International Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
criteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing, and some that are legal are
far from ISAF definition of optimality.

@@

Criteron~6 allows us to measure the \emph{balance} of a schedule. Perfect balance would be when a skipper has as many matches in first position in flights as in second position in flights, and so on. For example, if we had 9 skippers and 8 boats, each skipper would have 8 matches and could be in the first, second, third or fourth match of a flight. In perfect balance, each of the 8 skippers would appear twice in each position. Balance is one of our optimisation criteria.

@@

        \begin{scope}
            \node [anchor=north west] at (7.5, 0.3) {
                \begin{minipage}{4.5cm}
                    \centering
                    \setlength{\tabcolsep}{3pt}
                    \begin{tabular}{cccc}
                        \toprule
                        Flight & \multicolumn{3}{c}{Matches} \\ \midrule
                        0 & (0,6) & (3,2) & (5,4) \\
                        1 & (0,2) & (4,1) & (6,5) \\
                        2 & (2,1) & (5,0) & (6,3) \\
                        3 & (1,5) & (3,0) & (4,6) \\
                        4 & (3,5) & (6,1) & (2,4) \\
                        5 & (1,3) & (0,4) & (5,2) \\
                        6 & (4,3) & (2,6) & (1,0) \\ \bottomrule
                    \end{tabular}
                \end{minipage}
            };
        \end{scope}

@@

We present four pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by
ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF
manual, for 9 skippers and 6 boats and 9 skippers with 8 boats, then 10 skippers and
6 boats, again ISAF published and generated by us. Finalyy, we show what is a relatively easy problem, 8 skippers and 8 boats. 
For each of our schedules (with the exception of the "easy-8" problem), we used 1 day CPU time for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3 and 4 (renaming and orientation) completed in less than a second.

@@

\begin{center}
\begin{figure}[h]
\begin{minipage}[t]{0.48\textwidth}
    \centering
\begin{tabular}{ccccc}
        \toprule
        Flight & \multicolumn{4}{c}{Matches} \\ \midrule
0 & (5,2) & (4,3) & (1,6) & (0,7) \\
1 & (2,4) & (0,6) & (1,7) & (5,3) \\
2 & (0,5) & (7,2) & (6,3) & (4,1) \\
3 & (7,3) & (5,1) & (4,0) & (6,2) \\
4 & (3,0) & (4,7) & (6,5) & (2,1) \\
5 & (6,4) & (7,5) & (2,0) & (3,1) \\
6 & (7,6) & (5,4) & (3,2) & (1,0) \\ \bottomrule
    \end{tabular}
\label{08-08a}
\end{minipage}
\hfill
\begin{minipage}[t]{0.48\textwidth}
    \centering
\begin{tabular}{ccccc}
        \toprule
        Flight & \multicolumn{4}{c}{Matches} \\ \midrule
0 & (4,3) & (2,0) & (1,5) & (7,6) \\
1 & (3,2) & (0,4) & (6,1) & (5,7) \\ 
2 & (4,1) & (7,3) & (5,2) & (0,6) \\
3 & (1,7) & (6,5) & (3,0) & (2,4) \\ 
4 & (0,5) & (4,6) & (7,2) & (3,1) \\
5 & (7,0) & (2,1) & (6,3) & (5,4) \\ 
6 & (6,2) & (5,3) & (4,7) & (1,0) \\ \bottomrule
    \end{tabular}
\label{08-08b}
\end{minipage}
\caption{Schedules for 8 skippers and 8 boats. On the left, the published ISAF schedule (illegal),
and on the right, our schedule.}
\label{08-08}
\end{figure}
\end{center}

\paragraph{8 skippers and 8 boats (easy-8):} Table \ref{08-08} shows the ISAF schedule on the right for 8
skippers and 8 boats and on the right, our schedule.  The ISAF schedule violates criterion~4 (last match in a flight cannot be first in next flight) for skipper 5 in flights 1 and 2. Also, criterion 5 (no skipper should have more than 2 consecutive port or starboard assignments) is violated for skippers 0, 1, 3, 4 and 6. Furthermore, skipper 1 appears in the last match of four flights resulting in significant imbalance. Our schedule, on the right, satisfies all criteria, is optimal (maximum imbalance of 1) and took less than 10 seconds in total (all stages) to produce.

>>>>

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e.\ \emph{skippers}, compete against each other in a series of matches taking
place in rounds, known as \emph{flights}.  A match is composed of two skippers, with each skipper in
a boat on their own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the International Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
criteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing, and some that are legal are
far from optimal.

@@

Criteron~6 allows us to measure the \emph{balance} of a schedule. Perfect balance would be when a skipper has as many matches in first position in flights as in second position in flights, and so on. For example, if we had 9 skippers and 8 boats, each skipper would have 8 matches and could be in the first, second, third or fourth match of a flight. In perfect balance, each of the 8 skippers would appear twice in each position.

@@

        \begin{scope}
            \node [anchor=north west] at (7.5, 0.3) {
                \begin{minipage}{4.5cm}
                    \centering
                    \setlength{\tabcolsep}{3pt}
                    \begin{tabular}{cccc}
                        \toprule
                        Flight & \multicolumn{3}{c}{Matches} \\[2pt] \midrule
                        0 & (0,6) & (3,2) & (5,4) \\
                        1 & (0,2) & (4,1) & (6,5) \\
                        2 & (2,1) & (5,0) & (6,3) \\
                        3 & (1,5) & (3,0) & (4,6) \\
                        4 & (3,5) & (6,1) & (2,4) \\
                        5 & (1,3) & (0,4) & (5,2) \\
                        6 & (4,3) & (2,6) & (1,0) \\ \bottomrule
                    \end{tabular}
                \end{minipage}
            };
        \end{scope}

@@

We present three pairs of schedules. The first pair is for 8 skippers and 6 boats, one published by
ISAF, the other generated by us. Next, we present two new schedules, not appearing in the ISAF
manual, for 9 skippers and 6 boats and 9 skippers with 8 boats. The last pair is for 10 skippers and
6 boats, again ISAF published and generated by us. For each of our schedules, we used 1 day CPU time
for the first stage (minimising boat changes) then a further 2 days to minimise imbalance. Stages 3
and 4 (renaming and orientation) completed in less than a second.

@@
